---
title: "Project VAR"
author: "Chris Reimann"
date: "11/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd('/Users/chrisreimann/Desktop/U_Chicago/MSCA_31006_Time_Series/Project')
```

```{r}
library(expsmooth)
library(knitr)
library(tidyverse)
library(tseries)
library(TSA)
library(xts)
library(lubridate)
library(timetk)
```


```{r}
sales <- read_csv("MonthlySales.csv")
Chicago_sales <- sales[3,6:length(sales)]
Chicago_sales <- Chicago_sales %>% gather(key = "Date", value = "Price")
Chicago_sales$Date<-as.POSIXct(Chicago_sales$Date, format="%m/%d/%y")
Chicago_TS<-ts(Chicago_sales$Price, start = c(2008,2), frequency = 12)
Chicago_TS
autoplot(Chicago_TS)
```

```{r}
Chicago_train<-window(Chicago_TS, end = c(2018,12))
Chicago_test<-window(Chicago_TS,start = c(2019,1), end = c(2019,12))
```


#Load potential variables
```{r}
CPI <- read_csv("CPImonthly.csv")
CPI$Date<-as.POSIXct(CPI$Date, format="%m/%d/%y")

mortrate <- read_csv("mortgagerate.csv")
mortrate$Date<-as.POSIXct(CPI$Date, format="%m/%d/%y")

UnempR <- read_csv("UnemploymentRate.csv")
UnempR<-UnempR[1:151,]
UnempR$Date<-as.POSIXct(UnempR$Date, format="%m/%d/%y")

DJI <- read_csv("DJI.csv")
DJI$Date<-as.POSIXct(DJI$Date, format="%m/%d/%y")

econ <- read_csv("Chicago_econ.csv")
econ$Date<-as.POSIXct(econ$Date, format="%m/%d/%y")

CS_Index <- read_csv("CHXRNSA.csv")
CS_Index$Date<-as.POSIXct(CS_Index$DATE, format="%m/%d/%y")
```

```{r}
#Window
train_end<-c(2018,12)
test_start<-c(2019,1)
test_end<-c(2019,12)

Chicago_train<-window(Chicago_TS, end = train_end)
Chicago_test<-window(Chicago_TS,start = test_start,end = test_end)

CPI_ts<-ts(CPI$CPI, start = c(2008,2), frequency = 12)
CPI_train<-window(CPI_ts, end = train_end)
CPI_test<-window(CPI_ts,start = test_start,end = test_end)

mortgage_ts<-ts(mortrate$Rate, start = c(2008,2), frequency = 12)
mortgage_train<-window(mortgage_ts, end = train_end)
mortgage_test<-window(mortgage_ts,start = test_start,end = test_end)

Unemp_ts<-ts(UnempR$UnempRate, start = c(2008,2), frequency = 12)
Unemp_train<-window(Unemp_ts, end = train_end)
Unemp_test<-window(Unemp_ts,start = test_start,end = test_end)

DJI_ts<-ts(DJI$Close, start = c(2008,2), frequency = 12)
DJI_train<-window(DJI_ts, end = train_end)
DJI_test<-window(DJI_ts,start = test_start,end = test_end)

emp_ts<-ts(econ$employment, start = c(2008,2), frequency = 12)
emp_train<-window(emp_ts, end = train_end)
emp_test<-window(emp_ts,start = test_start,end = test_end)

CSI_ts<-ts(CS_Index$Index, start = c(2008,2), frequency = 12)
CSI_train<-window(CSI_ts, end = train_end)
CSI_test<-window(CSI_ts,start = test_start,end = test_end)


cor(CPI_train,Chicago_train)
cor(mortgage_train,Chicago_train)
cor(Unemp_train,Chicago_train)
cor(emp_train,Chicago_train)
cor(DJI_train,Chicago_train)
cor(CSI_train,Chicago_train)
```

#Select top 3 correlated variables: CSI Index, Employment, and mortgage rate.
```{r}
library(vars)
VARselect(cbind(Chicago_train,CSI_train,emp_train,mortgage_train), lag.max = 20, type = c("const", "trend", "both", "none"),
season = 12, exogen = NULL)$selection
```

```{r}
var1<-VAR(cbind(Chicago_train,CSI_train,emp_train,mortgage_train),p=7,type = "both", season = 12)#We will not use mortgage rate as it is the least coorelated. Effort to keep Ks to a minimum, 
serial.test(var1, lags.pt = 20, type = "PT.asymptotic") 
```
### Best p-value is 7, but we reject null hypothesis. Which means there is still some serial correlation.

#Select top 3 correlated variables
```{r}
VARselect(cbind(Chicago_train,CSI_train,emp_train), lag.max = 20, type = c("const", "trend", "both", "none"),
season = 12, exogen = NULL)$selection
```

```{r}
var2<-VAR(cbind(Chicago_train,CSI_train,emp_train),p=7,type = "both", season = 12)#We will not use mortgage rate as it is the least coorelated. Effort to keep Ks to a minimum, 
serial.test(var2, lags.pt = 20, type = "PT.asymptotic") 
```
### We are able to accept the null hypothesis, no serial correlation.


#Select top 3 correlated variables
```{r}
library(vars)
VARselect(cbind(Chicago_train,CSI_train,mortgage_train), lag.max = 20, type = c("const", "trend", "both", "none"),
season = 12, exogen = NULL)$selection
```

```{r}
var3<-VAR(cbind(Chicago_train,CSI_train,mortgage_train),p=5,type = "both", season = 12)#We will not use mortgage rate as it is the least coorelated. Effort to keep Ks to a minimum, 
serial.test(var3, lags.pt = 20, type = "PT.asymptotic") 
```
### Best p-value is 7, but we reject null hypothesis. Which means there is still some serial correlation.


```{r}
acf(residuals(var1)[,1])
acf(residuals(var2)[,1])
acf(residuals(var3)[,1])
```
### Our second model has the best looking ACF graph.
```{r}
summary(var2)
```
## NEED TO ADD ANALYSIS OF MODEL

```{r}
var_frcst1<-ts(predict(var1, n.ahead = 12, ci = .95)$fcst$Chicago_train[,1],start = test_start, frequency = 12)
var_frcst2<-ts(predict(var2, n.ahead = 12, ci = .95)$fcst$Chicago_train[,1],start = test_start, frequency = 12)
var_frcst3<-ts(predict(var3, n.ahead = 12, ci = .95)$fcst$Chicago_train[,1],start = test_start, frequency = 12)

accuracy(var_frcst1,Chicago_test)
accuracy(var_frcst2,Chicago_test)
accuracy(var_frcst3,Chicago_test)
```

### Our second model performs best and does not show serial autocorrelation and has a great acf plot. We will perform CV on this model.

#CV
```{r}
k <- 72 # minimum observations (6 years)
p <- 12 # Period
st <- 2014+11/12
```


```{r}
error <- matrix(NA,36, 12)
dates <- c()
for(i in (1:36))
  
{
  train <- window(Chicago_TS, start=st+(i-k+1)/p, end=st+i/p)
  test <- window(Chicago_TS, start=st + (i+1)/p, end=st+(i+12)/p)
  CSI_train <- window(CSI_ts, start=st+(i-k+1)/p, end=st+i/p)
  CSI_test <- window(CSI_ts, start=st + (i+1)/p, end=st+(i+12)/p)
  emp_train <- window(emp_ts, start=st+(i-k+1)/p, end=st+i/p)
  emp_test <- window(emp_ts, start=st + (i+1)/p, end=st+(i+12)/p)
  dates[i] <- st+(i+12)/p
  var_model <-  VAR(cbind(train,CSI_train,emp_train),p=7,type = "both", season = 12)
  
  f <- ts(predict(var_model, n.ahead = 12, ci = .95)$fcst$train[,1],start =st + (i+1)/p, frequency = 12)
  error[i,1:length(test)] <- f-test
}

MAE <- colMeans(abs(error), na.rm=TRUE)
RMSE <- sqrt(colMeans(error**2, na.rm=TRUE))
plot(1:12, RMSE, type="l",col=3,xlab="Horizon", ylab="Error",
     main='Cross Validation Results: 1/31/2016 to 12/31/2018') 
lines(1:12, MAE, type="l",col=2)
legend("topright",legend=c("RMSE", "MAE"),col=1:4,lty=1)
print(MAE)
print(RMSE)
print(sum(MAE)/12)
print(sum(RMSE)/12)
plot(dates, error[,12], ylab='Residual', main='Residuals from 12 Month Horizon')
```

### Interestingly, we are able to predict better at 12 months than 6 months. 


## Test Set Results
```{r}
error <- matrix(NA,12, 12)
dates <- c()
for(i in (37:48))
  
{
  train <- window(Chicago_TS, start=st+(i-k+1)/p, end=st+i/p)
  test <- window(Chicago_TS, start=st + (i+1)/p, end=st+(i+12)/p)
  CSI_train <- window(CSI_ts, start=st+(i-k+1)/p, end=st+i/p)
  CSI_test <- window(CSI_ts, start=st + (i+1)/p, end=st+(i+12)/p)
  emp_train <- window(emp_ts, start=st+(i-k+1)/p, end=st+i/p)
  emp_test <- window(emp_ts, start=st + (i+1)/p, end=st+(i+12)/p)
  dates[i-36] <- st+(i+12)/p
  final_md<-VAR(cbind(train,CSI_train,emp_train),p=5,type = "both", season = 12)
  
  f <- ts(predict(final_md, n.ahead = 12, ci = .95)$fcst$train[,1],start =st + (i+1)/p, frequency = 12)
  error[i-36,1:length(test)] <- f-test
}
MAE <- colMeans(abs(error), na.rm=TRUE)
RMSE <- sqrt(colMeans(error**2, na.rm=TRUE))
plot(1:12, RMSE, type="l",col=3,xlab="Horizon", ylab="Error",
     main='Test Set Results: 1/31/2019 to 12/31/2019') 
lines(1:12, MAE, type="l",col=2)
legend("topright",legend=c("RMSE", "MAE"),col=1:4,lty=1)
print(MAE)
print(RMSE)
print(sum(MAE)/12)
print(sum(RMSE)/12)
plot(dates, error[,12], ylab='Residual', main='Residuals from 12 Month Horizon')
plot(f)
points(test)
```
